---
title: "Untitled"
output: html_document
date: "2024-12-16"
---
# Purpose:
# Convert spatial transcriptomics objects between formats for downstream use:
# - SpatialExperiment <-> AnnData (via zellkonverter/reticulate)
# - SpatialExperiment -> Zarr
# - Seurat -> AnnData (via SeuratDisk)

# ------------------------------------------------------------------------------
# Inputs
# ------------------------------------------------------------------------------
# - spe_filt (SpatialExperiment object in R session)
# - 20250204_nnSVG.rds (SpatialExperiment with nnSVG results)
# - seurat_nnSVG.rds (Seurat object)

# ------------------------------------------------------------------------------
# Outputs
# ------------------------------------------------------------------------------
# - AnnData object in Python (`adata`) and back-converted SCE (`converted`)
# - output_nnSVG.zarr (Zarr store of nnSVG SpatialExperiment)
# - seurat_nnSVG.h5ad (AnnData file from Seurat object)
# ------------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r convert spatial experiment object to anndata object}
#devtools::install_github("https://github.com/theislab/zellkonverter")

#These functions expect that r CRANpkg("reticulate") has already been loaded along with an appropriate version of the anndata package
#from terminal, in r-reticulate environment, I installed anndata with pip install anndata
reticulate::py_install(c("anndata", "scanpy"))
#convert spatial object loaded in R to anndata object
adata <- zellkonverter::SCE2AnnData(spe_filt)
 
# Maybe do some work in Python on 'adata':
    # BLAH BLAH BLAH

    # Convert back to an SCE:
#converted <- zellkonverter::AnnData2SCE(adata)

#load and convert spatial or single cell object 
#or read and convert h5ad adata object to rds
singlecell <- zellkonverter::readH5AD("/run/user/1804238067/gvfs/sftp:host=clust1-sub-1,user=lythgo02/mnt/scratchc/fmlab/lythgo02/OV_visium/emily/cell2location/upk10_sc_400_newlyAnnot.h5ad")
# Convert SCE to Seurat
seurat_obj <- as.Seurat(
  singlecell,              # your SingleCellExperiment object
  counts = "counts", # assay to use as raw counts
  data = "logcounts" # optional: assay to use as normalized data
)

# Save the Seurat object (R-native, metadata preserved)

saveRDS(singlecell, file="/tmp/upk10_sc_400_newlyAnnot_noconvert.rds")
saveRDS(seurat_obj, file="/run/user/1804238067/gvfs/sftp:host=clust1-sub-1,user=lythgo02/mnt/scratchc/fmlab/lythgo02/OV_visium/emily/cell2location/upk10_sc_400_newlyAnnot.rds")

adata <- zellkonverter::readH5AD("/run/user/1804238067/gvfs/sftp:host=clust1-sub-1,user=lythgo02/mnt/scratchc/fmlab/lythgo02/OV_visium/emily/cell2location/final_adata_vis_clean.h5ad")
# Convert SCE to Seurat
seurat_obj <- as.Seurat(
  adata,              # your SingleCellExperiment object
  counts = "raw_counts", # assay to use as raw counts
  data = "X" # optional: assay to use as normalized data
)

#NOTE SEE BELOW TO PRESERVE ALL OF THE INFO IN THE SPATIAL OBJECT INCLUDING COORDINATE AND IMAGE DETAILS 
# Save the Seurat object (R-native, metadata preserved)
saveRDS(adata, file="/tmp/final_adata_vis_clean_noconvert.rds")

```
```{r}

library(zellkonverter)
library(Seurat)
library(SeuratObject)
library(dplyr)

# 1. Load the AnnData object
adata <- zellkonverter::readH5AD(
  "/run/user/1804238067/gvfs/sftp:host=clust1-sub-1,user=lythgo02/mnt/scratchc/fmlab/lythgo02/OV_visium/emily/cell2location/final_adata_vis_clean.h5ad"
)

# 2. Convert to Seurat (expression data only)
seurat_obj <- as.Seurat(
  adata,
  counts = "raw_counts",
  data = "X"
)

# 3. Extract sample per spot
sample <- colData(adata)$sample
seurat_obj@meta.data$sample <- sample

# 4. Extract coordinates and image info
coords <- reducedDim(adata, "spatial")
coords_list <- split(as.data.frame(coords), sample)
image_list <- metadata(adata)$spatial


# 5. Assign images and coordinates to Seurat in a CellTrek-compatible way
for (sample_name in names(coords_list)) {
  sp_coords <- coords_list[[sample_name]]
  img_info <- image_list[[sample_name]]
  
  # Make sure image is an array (CellTrek expects array)
  seurat_obj@images[[sample_name]] <- list(
    image = as.array(img_info$image),
    scale.factors = img_info$scalef,
    coordinates = sp_coords
  )
}
# 5. Store original coordinates for reference
seurat_obj@misc$spatial_coords <- coords_list
seurat_obj@misc$spatial_images <- image_list

# 8. Save Seurat object
saveRDS(seurat_obj, file="/tmp/final_adata_vis_clean_seurat.rds")

```

```{r}
# Convert SpatialExperiment to Zarr (R)
library(SpatialExperiment)
library(zarr)

# Load nnSVG results
spe_nnSVG <- readRDS("20250204_nnSVG.rds")

# Convert to Zarr format
zarr::save_zarr(spe_nnSVG, "output_nnSVG.zarr")

# Convert Seurat to AnnData (R)
library(SeuratDisk)

# Convert Seurat object to AnnData (H5AD format)
seurat_obj <- readRDS("seurat_nnSVG.rds")
SaveH5Seurat(seurat_obj, filename = "seurat_nnSVG.h5Seurat")
Convert("seurat_nnSVG.h5Seurat", dest = "h5ad", overwrite = TRUE)

#to convert seurat object to h5ad, first convert to sce, then h5ad 
library(SingleCellExperiment)
sce <- as.SingleCellExperiment(test)

zellkonverter::writeH5AD(
    sce,
    file = "merged_seurat_upk10sc.h5ad")
```


```{r}
library(zellkonverter)  # for readH5AD
library(Seurat)

# Folder where you saved your split .h5ad files
split_dir <- "/home/lythgo02/Documents/OV_visium/split_by_sample/"  # <-- adjust
files <- list.files(split_dir,
                    pattern = "^final_adata_vis_clean_\\d+\\.h5ad$",
                    full.names = TRUE)

# Sort for reproducibility (1..6)
files <- files[order(nchar(files), files)]

# Read each file and convert to Seurat
seurat_list <- lapply(files, function(f) {
  adata <- zellkonverter::readH5AD(f)
  seurat_obj <- as.Seurat(
    adata,
    counts = "raw_counts",
    data   = "X"
  )
  return(seurat_obj)
})

# Give them names sample_1 … sample_6
names(seurat_list) <- sub("^.*_(\\d+)\\.h5ad$", "sample_\\1", basename(files))

# Quick sanity check
sapply(seurat_list, dim)

# Save each Seurat as RDS (optional)
out_rds_dir <- file.path(split_dir, "rds")
dir.create(out_rds_dir, showWarnings = FALSE)

invisible(lapply(names(seurat_list), function(nm) {
  saveRDS(seurat_list[[nm]], file = file.path(out_rds_dir, paste0(nm, ".rds")))
}))

# Or save the whole list as one RDS
saveRDS(seurat_list, file = file.path(out_rds_dir, "seurat_list_all.rds"))

```

